\documentclass[../../main.tex]{subfiles}
 
\begin{document}
\label{sec:teaching}

I have reflected on my overall teaching practice for the academic year (AY) of 2021-2022.  In Sec. \ref{sec:teaching_philosophy}, I respond to your questions in our last communication regarding my teaching philosophy.  In Sec. \ref{sec:intro}, I reflect on the introductory physics courses I taught in AY 2021-2022.  In Sec. \ref{sec:intro_eval}, I analyze the course evaluation data from these courses and reflect upon the student feedback.  In Sec. \ref{sec:adv}, I reflect on the advanced physics courses I taught in AY 2021-2022.  In Sec. \ref{sec:adv_eval}, I analyze the course evaluation data from these courses and reflect upon the student feedback.  I list several interesting ideas for modifying PHYS306/COSC330: Computer Logic and Digital Circuit Design, including further integration with COSC360: Digital Signal Processing.

\section{Teaching Philosophy: Six Easy Pieces of Active Learning Technique}
\label{sec:teaching_philosophy}

\begin{flushleft}
\subfile{teaching_philosophy/teaching_philosophy}
\end{flushleft}

\section{Introductory Course Descriptions}
\label{sec:intro}

\begin{flushleft}
\subfile{intro_course_desc/intro_course_desc}
\end{flushleft}

\section{Analysis of Course Evaluations: Introductory Courses}
\label{sec:intro_eval}

\begin{flushleft}
\subfile{intro_course_desc_eval/intro_course_desc_eval}
\end{flushleft}

\section{Advanced Course Descriptions}
\label{sec:adv}

\begin{flushleft}
\subfile{adv_course_desc/adv_course_desc}
\end{flushleft}

\section{Analysis of Course Evaluations: Advanced Courses}
\label{sec:adv_eval}

\begin{flushleft}
\subfile{adv_course_desc_eval/adv_course_desc_eval}
\end{flushleft}

\section{The Future}
\label{sec:future_teaching}

\begin{flushleft}
\subfile{future/future}
\end{flushleft}

\end{document}
