\documentclass[../../../main.tex]{subfiles}

\begin{document}
\label{sec:committee_service}

Here I reflect primarily on my service within the Educational Policy Committee.  Since I have been accepted onto the Whittier Scholars Program (WSP) Council, I also reflect on our plans to enhance WSP.  Finally, I offer some ideas on how I can serve the college in the future through institutional research.

\subsection{Educational Policy Committee}

Though the charge to EPC in AY 2021-2022 contained many projects and goals, none required as much energy and thorough deliberation as finishing the course system proposal.  I must be honest when I share my reflections on these experiences: they were challenging.  I trust you to read what I've written within the context of honest reflection and service rather than the eventual fate of the course system proposal.
\\
\vspace{0.25cm}
When I joined EPC, I learned that people have been proposing a change to a course system for 20 years.  I asked about the original reasoning, and why my colleagues had spent large portions of their careers advocating for it.  Having served at four institutions of higher education and three Catholic parishes, I realized this debate was similar to other entrenched debates I have observed at my previous institutions.  The original reasons have become blurred, those whose minds were once open have become more solidified, and new factors complicate the discussion.  Examples of the latter are the evolving curriculum and the faculty workload.  Though the workload is not strictly a pedagogical issue, it is so important to faculty at Whittier College that it had to be included.  Entrenched debates lead to large amounts of accumulated time spent in meetings discussing the proposal.  Time \textit{itself} eventually becomes part of the debate: ``It would be a shame not to pass this ... we've already spent so much time working on it.''  Given my experiences working at other institutions, these traits are the hallmarks of debate that has become entrenched.
\\
\vspace{0.25cm}
In his lecture entitled ``Two Sacred Incompatible Values at American Universities,'' \cite{haidt} Prof. Jonathan Haidt (NYU Stern School of Business) defines \textit{motivated reasoning} in the following terms.  When someone wants to believe something, they ask internally ``\textit{Can} I believe this?'' and they tend to search for a single fact to support the claim.  When someone does not want to believe something, they ask internally ``\textit{Must} I believe this?'' and they tend to search for one piece of evidence to reject it.  Psychological studies suggest that people engage in motivated reasoning when considering for/against propositions with moral or practical implications for their lives.  In our EPC discussions, the reasons why members \textit{wanted} or \textit{did not want} to believe the course system would benefit Whittier College eventually became clear.  Some wanted it because they felt it would help our finances.  Others wanted it because they felt it would help their advisees graduate given the loss of January term.  Some did not want it because they felt it would hurt a major degree within their department.  Some did not want it because it seemed to grant degrees to students who do not complete enough work.  
\\
\vspace{0.25cm}
No matter what our internally motivated reasons were, we had to frame our discussions pedagogically as we were instructed.  One of our tasks was to define a standard course.  Courses can be as diverse as human thought: some are lighter, some more intense, and some are more narrow or broadly focused.  The goal was a solid course definition, and consensus for the total number of \textit{courses} required to graduate.  Motivated reasoning leads us to different results, but justifying the numbers pedagogically is hard.  We were instructed to use 32 courses as the standard.  With a total of 32, a student could take courses that corresponded formerly to 3 credit-hours for a degree with 96 credit-hours, compared to the original 120 to graduate under the credit system.  The same calculation with 4 credit courses leads to 128 credits, and performing a weighted average of the extremes still produces results that are lower than we currently require.  The ratio of 3 credit courses to 4 credit courses depends on the student, and we examined those statistics as well.  We actually compared the workload for 3 and 4 credit courses.  I had assummed a 4 credit course is more work, but I learned the importance of defining and measuring work outside of class.  I found that most of us were relying on personal experiences to judge the workload outside of class.
\\
\vspace{0.25cm}
In response to the difficulty of developing consensus around courses, we took several logical steps to ensure students will graduate smoothly and on time under the credit system.  We raised the students' per semester credit limit to 17, and worked together to secure the approval of the faculty.  We also updated the definition of a credit-hour, after reviewing material from 4-year liberal arts colleges within and outside of WASC regarding total credits to graduation given their credit-hour definitions.  We began a sub-committee project, which I led, to understand and classify ``tracks,'' ``emphases,'' and ``options,'' within department majors.  The goal for that sub-committee project is to standardize language surrounding major programs in order to help students understand their options and plan for graduation.  I reflect on the outcome of this project below, after I share my reflection on the course system proposal.
\\
\vspace{0.25cm}
Another component of the proposal was the proposed cap on courses per major.  Since a bedrock goal of the system is graduation in 4 years, there is some upper limit on the number of courses even the most intrepid students can take.  With a finite number of courses to select, it would be unexpected for a student at a liberal arts college to chose 80 \% of their classes from one subject.  Some of us were pressing for a cap of 11-12 courses, with the thought being one third of the courses chosen for the major, one third core liberal arts requirements, and one third for ``exploration.''  The counter-argument is very simple: some major programs cannot exist with 11-12 courses.  Consider the KNS major with ``pre-physical therapy emphasis''\footnote{In 2021, 10.5\% of all students chose this major.  Compare this number to 14\% for BSAD.}.  I can imagine two ways to estimate the number of courses in the major: count the courses currently required, or total the credits and divide by an appropriate denominator.  The KNS department lists 21 courses required to graduate with this emphasis.  We also find about 21 courses if we total the credits and divide by 3.5 credits per course\footnote{This major contains 14 courses worth 4 credits each.}.
\\
\vspace{0.25cm}
Thus, capping majors at 11-12 courses would eliminate one of our most popular majors.  Our discussions did yield a potential compromise: counting only courses beginning with KNS (that is, the department of the major) towards the cap.  The total KNS courses in the major would stay under the cap (11), since the other courses come from CHEM, BIOL, MATH, PHYS, and PSYC.  This does not guarantee that students would take more courses in the humanities and social sciences outside of PSYC and KNS. Through our EPC discussions, it became clear to me that the motivated reasoning behind the cap on major courses was to encourage students to take courses in the humanities and social sciences that need more students.  I consider my conscience well-formed on this point, and I cannot ignore my own moral instincts.  While a central facet of a liberal arts college is that students receive a well-rounded education, we cannot make students take courses they don't wish to take.  The moral responsibility for course selection lies with the students and their families.  That being said, some appropriate cap must always be put in place, and even the current system has a credit-limit for majors.
\\
\vspace{0.25cm}
Consider one more example: ICS/Mathematics.  ICS/Math has experienced 950\% growth since 2018, and it has an average female-to-male ratio of 44\%.  The popularity and diversity of this major at Whittier College is an asset to us, and students are often unable to enroll in similar programs at larger schools due to overcrowding.  We have an opportunity to expand our undergraduate base by meeting the needs of students.  However, this major has 17 courses.  The compromise we devised above does not apply, for this major is interdisciplinary between math and computer science and therefore all 17 courses count towards the major.  Enforcing a cap to encourage students to take other courses would significantly alter one of our fastest growing majors.  I also note that a female-to-male ratio of 44\% is more than double the result for all 3-2 Engineering majors over the last six years.  Thus, we should cherish and build on such diversity in the ICS/Math Program.  Thus, I've shown how including a major cap as part of the course system proposal would have negatively impacted one of our most popular majors (KNS with pre-physical therapy), and one of our fastest growing ones (ICS/Math).
\\
\vspace{0.25cm}
At one point, the discussion of the course system proposal was steered back to financial reasoning.  We reviewed a series of calculations that demonstrated a course system with 32 courses to graduate would save financial resources.  The mechanism appeared to be that the average courses per student per semester would decline, with tuition remaining constant.  What we found strange was that we were initially instructed to focus on strictly pedagogical reasoning for the course system proposal.  I agree that there was a modest financial advantage to the course system over credit, but it arises because on average students are taking fewer courses under the specific scenario we examined.  Some of us felt that reducing the courses per student per semester was akin to granting a degree for less work\footnote{Alternatively, we can think of this as indirectly raising tuition by collecting the same tuition for fewer courses.  These are statistical statements drawn from the scenario we examined, which focused on averaging over hundreds of students.}.  Once again, these are my honest reflections and I do not mean to suggest that my colleagues wanted to do the wrong thing.  We all had good intentions, and we tried really hard to understand the effects a complex issue would have on our students.
\\
\vspace{0.25cm}
Towards the end of the year we resolved the impasse regarding the cap on the number of courses per semester students could take.  The original proposal blocked first-year students from taking five courses in Fall or Spring, and any student would have had to obtain special permission to take five courses.  I understand the reason is that we are trying to protect students from failing out of college.  Given my experience advising in 3-2 Engineering, Physics, and ICS students, I believe course caps will lead to painful setbacks to graduation.  On EPC, we examined articles covering data from other institutions that suggests a strong load is positively correlated with student success.  We also noticed in our own data that students who take five courses tend to have higher grades, but that only $\approx 10$\% of the students take five courses or 16-17 credits.  Two camps developed: no per-semester caps at all, or the original proposal.  I offered a compromise: during the first semester, a cap of four courses, followed by the ability to apply for and acquire the ability to take five courses in any semester given an acceptable GPA.  Academic probation would remove the ability.  Thus, we can measure a student's first-semester college GPA.  My ESAC committee studies indicate a modest GPA decrease in the transition from high school to college (Sec. 4.1.1 of my prior PEGP).  The compromise also respects the will of the students.  My solution satisfied both sides and was adopted into the proposal.
\\
\vspace{0.25cm}
Finally, I mentioned above that I led a sub-committee to study the partioning of major programs through ``options,'' ``tracks,'' and ``emphases.''  I gathered data from Whittier College websites and DegreeWorks, and compiled an executive summary (see Supplemental Material).  I chose the term ``partition'' as a neutral term for any sub-classification of a major program.  I found that the average partitions per major across 27 majors Whittier College is $2.2 \pm 1.3$.  From public-facing website data, the divisional breakdown was $3.3 \pm 1.0$, $1.8 \pm 1.3$, and $1.9 \pm 1.2$ for the physical sciences, social sciences, and humanities, respectively (mean and standard deviation).  The numbers were similar for data taken from DegreeWorks: $2.4 \pm 1.5$, $2.0 \pm 1.3$, and $1.9 \pm 1.1$ for physical sciences, social sciences, and humanities, respectively (mean and standard deviation).  The important conclusion to draw from the data is that very few departments have just a single version of the major discipline.  Terms like track or emphasis are used to partition majors into sub-classified programs.  EPC considered that students could better understand the sytem if terms were standardized.  This is especially true for first-generation students.  If similar terminology and structures were common throughout the major programs within an overall curricular design, we might find that students take better control of their educational planning.  For now, we have sent a departmental survey to department chairs, based on the preliminary results.  We hope the survey will help us better understand the usage of these terms and why majors are partitioned the way they are.

\subsection{The Whittier Scholars Program}

As I shared in my last PEGP report, I have already graduated one student through the Whittier Scholars Program.  This was a fantastic interdisciplinary experience.  Nicolas Bakken-French was an intrepid sophomore when he introduced himself to me, and he was interested in glaciers.  He had heard that at least part of my research involved Antarcticic science, and was curious if this connected to his plans.  He wanted to understand and document the disappearance of polar glaciers from a scientific and cultural perspective.  Our work sent him all over the world collecting data, gaining field experience and survival training, and photographs of ice fields and glaciers.  I was proud of the scientific and cultural report that we submitted in the end, and I felt that we produced a polished product.
\\
\vspace{0.25cm}
I have a new student in the Whittier Scholars Program, Jackson Diamond, who focuses on computer programming applications in medicine.  Jackson has been developing his innovation with me that is central to his final WSP project.  He is creating a position location system based on Bluetooth signal strengths and trilateration.  This would be useful for locating objects with Bluetooth tags with a precision of a few centimeters within a room or floor of a building.  Jackson shares that his first appliction of this technology would be patient tracking within a hospital.  Imagine a Bluetooth tag connected to a baby crib within a maternity ward that is integrated within a trilateration system.  The location of the baby could be remotely monitored for safety and security purposes.  I agreed to take on Jackson's program after his original advisor (Bill Kronholm) had to depart.  For me it was an easy service decision to make.  WSP needs new professors to step forward to maintain it, and it is a special privilege that we have such an interesting interdisciplinary program at our institution.
\\
\vspace{0.25cm}
Students like Jackson and Nicolas are why I have volunteered to serve on the WSP Council.  With Prof. Chihara now leading the council, our task this year is to streamline and update the program.  I look forward to working with Prof. Chihara by recruiting for WSP and enhancing the proposed interaction between INTD100 and WSP.  The basic idea is that some INTD100 sections would be integrated with introductory-level WSP courses for students who know they want to pursue interdisciplinary studies.  For my part, I will continue to serve as an INTD100 instructor over the years, and help strengthen the bond between the first-year experience and an introduction to WSP.  The goal, as Prof. Chihara says, is not necessarily to draw more students into WSP, but to show them the interdisciplinary nature of academia such that they choose a path for their major they find intriguing and impactful.  As part of this overall plan, I hope to strengthen ties between computer science as an interdisciplinary field and WSP.  Students like Jackson and Nicolas have shown me that there is still a hunger for a diverse liberal arts education among students in the physical sciences.  Through my advising and service, I hope my students and I will generate new ideas to serve our community.

\subsection{Future Proposals for Institutional Research}

In the various service roles I have taken in my time at Whittier College, I have noticed a common theme.  I have enough skill with data analysis, programming, and statistics to perform institutional research.  I have also noticed that some of our institutional research staff have recently retired.  For these reasons, I would like to suggest a few ideas for ways I could serve Whittier College in the future.
\begin{enumerate}
\item In my writing, I often make use of the data from the dashboards created by Gary Wisenand that are available on Moodle.  Using these data, for example, I determined the growth rate and gender breakdown of the ICS/Math program.  These dashboards need to be updated with data from AY 2021-2022 and beyond.  According to Gary, I would need to learn to use software tools like Tableau.  This would not be difficult given my experience level.  I am interested in questions like the demographic breakdown of introductory STEM courses, the participation of young women in STEM majors, and measuring the growth of the wide variety of major programs and the corresponding ``partitions.''
\item Since I served on ESAC for two years and EPC for one year (with ERC in between those), I have been exposed to a lot of analyses of student data.  An idea for an interesting statistical analysis has been floating around in the back of my mind, related to the probability of graduation.  To boost retention in graduation, I wonder if my colleagues would find it useful to characterize the well-worn and relatively uncommon pathways students take to graduate through our system.  It would be interesting to me as an educator to calculate and visualize the relative frequencies of the pathways students take throught 
\end{enumerate}

\end{document}