\documentclass[../../main.tex]{subfiles}
 
\begin{document}
\label{sec:scholarship}

We as facutly at Whittier College classify scholarship according to the Boyer model, and my scholarship tends to fall within the categories of the scholarship of discovery, and the scholarship of application.  In Sec. \ref{sec:neutrino} I describe my physics research in the area of ultra-high energy neutrino physics (UHE-$\nu$).  In Sec. 2 of my prior PEGP I shared how, through my scholarship, Whittier Colllege is now a member institution of the IceCube Collaboration (Gen2) (\url{https://icecube.wisc.edu}).  I have published original scholarly work with the help of an undergraduate research student that represents a significant contribution to the future of IceCube Gen2 (Sec. \ref{sec:neutrino}).  I also write about my future plans with high-performance computing (HPC) and machine learning applications for IceCube Gen2, and grant submissions we have written for this purpose.
\\
\vspace{0.25cm}
In Sec. \ref{sec:naval_research} I describe my growing relationship with NSWC Corona Division, a US Navy research laboratory in Corona, CA, through the Office of Naval Research (ONR).  Primarily an engineering and computer science research facility, NSWC Corona serves the Navy through advances in logistics, reliability research, metrology, and radio-frequency (RF) communications and radar.  My relationship with ONR and NSWC Corona has produced a publication, three summer fellowship grants, and grant resources like lab equipment for Whittier College.  The work has inspired collaboration with academics in engineering and physics at other institutions.  This summer I was an invited to speak about our ONR research at a national conference at the Massachusetts Institute of Technology (MIT).  Finally, in Sec. \ref{sec:naval_research} I am excited to share that our colleagues at NSWC Corona would like to form an Educational Partnership Agreement (EPA) with Whittier College.  If adopted, the EPA will provide enduring research experiences, internships, and career prospects in engineering for the next generation of Poets.

\begin{flushleft}
\subfile{neutrino/neutrino}
\end{flushleft}

\begin{flushleft}
\subfile{naval_research/naval_research}
\end{flushleft}

\end{document}
