\documentclass[../../main.tex]{subfiles}
 
\begin{document}

Dear Friends,
\\
\vspace{0.15cm}
I have compiled a report on my progress as an instructor, scholar, steward, and advisor for Whittier College during the period of 2021-2022.  The following is a reflection on the development of my educational and scholarly practices, and the service I have performed for the College as a mentor, advisor, and committee member. I strive to perfect my teaching abilities, and I am pleased to report that my students are learning and growing at Whittier, and achieving success in the professional world.  In our last communication, after my second major PEGP report (delayed to the fifth year due to the pandemic), you asked me to reflect on my pedagogical practices.  In particular, you suggested four concrete topics on which I could reflect.  First, you asked me to describe my interpretation of the learning process.  Second, you asked me to describe how I incorporate tools and practices in my courses.  Third, you asked me to describe how the tenets of my teaching philosophy help us to achieve the learning objectives that we set for our courses.  Finally, you asked me to focus on the \textit{why} behind specific teaching decisions as opposed to the \textit{how}.  You also posed similar questions about the \textit{learning focuses} that I have provided in the past.
\\
\vspace{0.25cm}
With these four simple questions in mind, I have taken what I think is the most straightforward and concise approach to the structure of my teaching philosophy in Sec. xxx.  I reflected on what practices I actually use most often in my teaching, and answered each of the four questions for all practices.  This exercise has been useful and enlightening, as it has encouraged me to think carefully about how the central principles of \textit{order} and \textit{shared meaning} are reflected in the subject of physics, and the instruction of physics.  It is my hope that this exercise provides you with useful insight into modern physics instruction.  I have also reflected on the \textit{learning focuses}.  These ideas were derived originally from my colleagues in my department, but I have since modified them and made them my own.  In reflecting on how I conduct my courses, I have come to the conclusion that the learning focuses simply guide my course creation and course content selection.  How I conduct my courses, and the \textit{why} behind specific teaching decisions I make are driven primarily by the tenets of my teaching philosophy.  Though my teaching practices continue to generate positive student feedback in the form of strong student evaluation scores, I have also identified areas of courses that need to be adjusted.
\\
\vspace{0.25cm}
Turning to my scholarship, I have many new and exciting accomplishments to share with you in Sec. xxx.  Three recent experiences come to mind as examples.  First, I have finally published in \textit{Physical Review D,} the flagship peer-reviewed journal in my field by the American Physical Society (APS)\footnote{J.C. Hanson and R. Hartig. ``Complex analysis of Askaryan radiation: a fully analytic model in the time domain.'' Phys. Rev. D \textbf{105}, 123019 (2022).}.  As far as I can tell, I am the first professor at Whittier College to achieve this.  This publication was the culmination of two years of work with an undergraduate student who has become a dear friend.  This result marks the first time a professor from my department has published in one of the \textit{Physical Review} journals in the last 16 years\footnote{See, for example, S. Zorba \textit{et al}. ``Fractal-mound growth of pentacene thin films.'' Phys. Rev. B \textbf{74}, 245410 (2006).}.  The piece provides the first fully analytic model of Askaryan radiation.  I hope to make clear why the results represent a significant contribution to my field, and how my undergraduate researcher helped me to improve and finish this work (see Sec. xxx, and also Sec. 3.3.2 of my previous PEGP).
\\
\vspace{0.25cm}
The second experience pertains to my radio-frequency (RF) engineering and radar research with the Office of Naval Research (ONR).  In 2021, I published a paper involving the computational electromagnetism (CEM) of radar design\footnote{J. C. Hanson.  ``Broadband RF Phased Array Design with MEEP: Comparisons to Array Theory in Two and Three Dimensions.'' Electronics Journal \textbf{10}, 415 (2021).}.  This work was ranked Top 10 Most Notable Articles in Electronics Journal for six months.  The work caught the attention of CEM experts from four different countries who each contacted me for advice and collaboration.  This Summer, I was invited to speak at a CEM conference held at MIT.  I gave a 45 minute lecture on open-source RF CEM design alongside colleagues from MIT, Google, Georgia Tech, Stanford, and BYU.  It was an incredibly meaningful moment in my career.  I took the opportunity to promote the mission and values of Whittier College to our peers at these institutions.  Having received an ONR Summer Faculty Research Program (SFRP) grant for the third year in a row, I am eligible for ONR Senior Fellowship in Summer 2024 after a mandatory one-year break.  For the past two years, our ONR partners at NSWC Corona Division have granted us money and precision RF equipment to boost the engineering research experiences of our students.  Based on this fruitful collaboration, I'm happy to share that they would like to form an Educational Partnership Agreement (EPA) with Whittier College.  NSWC Corona forms EPAs with colleges througout Southern California in order to strengthen undergraduate engineering research, and to recruit engineering talent.  Including Whittier College students in this endeavor will be wonderfully beneficial for our students.
\\
\vspace{0.25cm}
The third and final experience is related to both my scholarship and teaching.  In Fall 2019, I taught INTD255, entitled ``Safe Return Doubtful: History and Current Status of Modern Science in Antarctica'' (CON2).  In Spring 2021, I taught INTD290, entitled ``A History of Science in Latin America'' (CON2,CUL3).  I concluded these courses with material related to the connection between modern scientific endeavors by peoples of diverse cultures and exploration literature.  I showed the students how it is possible to travel to Antarctica through the United States Antarctic Program (USAP).  As part of my research with the IceCube Gen2 collaboration\footnote{See, for example, \url{https://icecube.wisc.edu/}.}, I have conducted research expeditions to Antarctica through USAP.  Thus, there is a deep connection between my teaching and research via the concept of exploring the unknown.  A long-time goal of mine has been to inspire my students to begin their own careers in science and in life with the same organization, mental discipline, curiosity, and confidence required of any explorer.  I am thrilled to report that we are finally sending our first Poet undergraduate to Antarctica.  Scout Mucher, who was my student in INTD290, was inspired to go, and we met to go over the application process in detail.  Scout was hired as a contractor to help run USAP operations in McMurdo station, our flagship base on Ross Island.  Scout has now PQ'd (physically qualified), and is slated to begin work there when the sun rises this Fall!  I am so proud of Scout, who will become the first Whittier Poet to set foot on Antarctic shores.
\\
\vspace{0.25cm}
For my service to Whittier College for the 2021-2022 academic year, I joined the Educational Policy Committee (EPC).  I had been in discussions with Prof. Rehn about joining the Whittier Scholars Program (WSP) advisory board.  Near the end of Spring 2021, I was called to first serve for a year on EPC, and I answered the call.  The major task for EPC was to generate consensus around a revised course system proposal to be brought before the full faculty by the end of the year, in light of proposed changes to faculty load.  While discussions of the course system and faculty load proposals continued, we also completed a long list of other tasks.  These included studying and approving changes to ten major programs and changes to programming within the Center for Engagement with Communities (CEC), raising student per-semester credit limits, revising the definition of a credit hour, revising the new course proposal form, revising handbook language pertaining to syllabi, and to study the use of ``tracks'' or ``emphases'' within major programs at Whittier College.
\\
\vspace{0.25cm}
I led a sub-committee dedicated to the study of tracks and emphases, and our goal was to develop a common language for tracks across campus.  I framed the task by comparing ``tracks'' and ``options'' to \textit{partitions}, as a computer hard drive is partitioned.  We examined data from Whittier College websites and DegreeWorks to accurately determine the number of partitions per program.  I found that, on average, major programs at Whittier College have two partitions per major. Rounded to integers, the natural sciences tend to have three partitions, while the social sciences and humanities tend to have two.  The variances, however, are large.  We used this diverse data set to formulate a survey we sent to department chairs.  The data set and chair responses will be used to formulate common policy for tracks and options within majors.  For example, students might find it useful to have a common language describing options within majors to improve understanding of the curriculum.
\\
\vspace{0.25cm}
With regards to the course system, I tried my very best to aid in the discussions by thinking through the technical implications of each tenet of the proposal.  One example of how I helped to advance the discussion was to offer a compromise between two positions regarding the maximum number of courses per semester.  One one hand, some wanted to limit first-year students to just four courses per semester, and require all students to obtain special permission to take five courses in any semester.  On the other hand, some objected to any course restriction.  I describe the compromise position I devised in Sec. xxx.  We ultimately adopted this compromise into the final proposal.  This episode in my committee service demonstrates that I do have the capability to build consensus, even when the task is technically challenging.  I give much credit to my colleagues on EPC, and especially co-chairs Profs. Camparo and Householder, for setting such a good example from which I can draw useful experience.  In Sec. xxx, I describe ways in which I might be of service to Whittier College in the future, given my reflections on my accumulated committee service.  Often, I find ways to aid with technical discussions by writing code and crunching numbers that add weight to our policy discussions.  Thus, I propose to help with institutional research, and provide a few ideas in Sec. xxx.
\\
\vspace{0.25cm}
In Sec. xxx I describe my accomplishments in advising and mentoring.  In our last communication, you gave concrete suggestions for my writing and reflection in this area.  Specifically, you asked me to expand on my philosophical approach to advising and mentoring.  You shared a concern that my advising leads students down a rigid path, and that creating digital profiles using services like LinkedIn narrows their outlook.  You also shared that you would prefer a self-reflective approach to the advising and mentorship section, rather than a step-by-step guide.  I have responded by reflecting on my advising and mentorship much more succintly than my previous PEGP report.  I describe how I employ a philosophical perspective that motivates my students to take charge of their future in a way that is useful for them and in alignment with their values and interests.  I have also included tangible results of successful student outcomes, such as internships, publications, and fellowships gained by my advisees.  I also discuss in Sec. xxx plans for taking first-year advisees again this year, and my new section of INTD100.  I think you will find my topic selection for my INTD100 advisees particularly useful given the current cultural climate.
\\
\vspace{0.25cm}
To your point about rigidity, I'm surprised that the counter-examples I provided in my last report did not ameliorate your initial reaction.  I do retain flexibility in my approach to advising, and I gave an example as proof.  I had an advisee who initially decided on majoring in physics, but realized in the midst of our discussions that his path lies in a different direction.  After appropriate reflection, we decided to change his major to Digital Art and Design (see Sec. 5.2.1 of my previous PEGP).  In Sec. xxx, I share how the decision tree I provided (also in Sec. 5.2 of my previous PEGP) is not just something I invented, but actually reflects the thinking of many majors in the physical sciences.  I have infused this decision tree with more detail about the diverse paths my students take towards their career goals.  I also point out in Sec. xxx that our recent curricular discussions include the creation of digital portfolios and encourage sharing senior theses in Poet Commons.  Our advisees will share these with potential employers, and use them to draw connections between projects at Whittier College and their future plans.  Tools liked LinkedIn are a widely accepted mechanism to facilitate sharing work.  It is fruitful to encourage students to use these tools to take control of their job search, and to refine and even reverse-engineer the search given their digital portfolio. Finally, I note that I will serve on the WSP advisory board this coming academic year.  My advising of current WSP students, WSP advisory board participation, and recruitment for the program all serve as evidence that my advising and mentorship is closer to the \textit{opposite} of rigid.  The very nature of WSP is to foster academic exploration in interdisciplinary and flexible settings.  In Sec. xxx, I describe my latest advising effort with my current WSP student, and future plans for new WSP students.
\\
\vspace{0.25cm}
In Sec. xxx, I describe a project regarding equity and inclusion within the natural sciences at Whittier College.  Back in 2017, I had the idea to create an app that would facilitate inclusion in introductory physics courses infused with machine learning to tailor the user's experience.  I named the idea ``The Primer,'' in reference to a sci-fi novel in which a young woman aquires a tool that accelerates her education through a narrative tailored to her background and past success in science.  The visual environment and narrative of the app was to be designed by diverse Whittier College undergraduates so as to make the experience as inclusive as possible.  I finally found time to encapsulate this idea into an internal DEI grant (submitted and approved by the Inclusion and Diversity Committee).  The IDC members suggested I attend three workshops on inclusivity in introductory STEM courses given by the Cottrell Scholars Network.  (Some of these workshops were later suggested to all departments by the Dean).  In these workshops, we learned about current psychological research within the field of inclusion in introductory STEM courses.  It is my hope that one of my current advisees in computer science, who comes from a diverse background, can help me create and test the code for this app over the coming year.  I will begin to recruit digital designers this Fall semester from within Whittier College to develop the digital storytelling aspect of the tool.  We hope everyone will enjoy the results!
\\
\vspace{0.25cm}
My friends, it is my hope to be granted tenure at Whittier College.  I hope that the effort and passion I have given to our institution for the past several years merits this distinction.  I have created and taught new courses that students have enjoyed and found useful, including liberal arts courses that fused my research interests in the natural sciences with social issues and history.  I have demonstrated tangible successes in the areas of scholarship of discovery and application, and the the future holds several exciting research avenues to explore with our students.  I have served on multiple committees that have tackled complex issues ranging from student admissions data to the course system proposal.  I have continued to be a resource as a first-year advisor to many new Whittier Poets, and I have a track record of leading majors in the physical science to successful career outcomes.  All the while, I have done my best to incorporate your suggestions made with professional candor into my work.  I speak for myself and my department when I say that we are grateful for the work that you do, and we trust in your insight and wisdom.
\\
\vspace{0.25cm}
Sincerely, \\
Jordan C. Hanson \\
Assistant Professor, Dept. of Physics and Astronomy
\end{document}
